---
title: "Spatial Statistics - Introduction"
author: "Adela Sobotkova"
date: "11/12/2020 updated `r format(Sys.time(), '%B %d, %Y')`" 
output:
  rmdformats::readthedown:
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("spatstat")
```
```{r}
library(spatstat)
```

# Task 7: Point Pattern testing

Two `ppp` objects have been created called `p1` and `p2` in your /data folder. Test if they are clustered or regular, and select the right answer. You'll need to run `quadrat.test` four times, once for each process with each hypothesis.

```{r}
ppp1 <- readRDS("../data/p1.rds")
ppp2 <- readRDS("../data/p2.rds")



# Run a quadrat test
quadrat.test(ppp1, alternative = "clustered")


# Run a quadrat test
quadrat.test(ppp1, alternative = "regular")


# Run a quadrat test
quadrat.test(ppp2, alternative = "clustered")


# Run a quadrat test
quadrat.test(ppp2, alternative = "regular")


```


# Task 3: Tree location pattern

Load a `ppp` object called `redoak` which contains the locations of trees in a woodland. Plot the tree locations and then test the hypothesis that the points are clustered against a null hypothesis of a random uniform distribution.

```{r, eval= FALSE}
redoak <- readRDS("../data/redoak.rds")

redoak_cluster_env <- envelope(redoak, correction = "border")

plot(redoak_cluster_env, . - pi * r ^ 2 ~ r)

```