---
title: "Working with sf objects"
author: "Adela Sobotkova"
date: "04-February-2022 updated `r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

#  Goals: 

1. In the first exercise, you will learn to construct maps with tmaps() by modifying the provided code. The libraries, data, and initial code is provided below.
2. In the second exercise, you learn to: 
 - Understand the provided datasets
 - Learn how to reproject spatial data
 - Limit your data into an area of interest
 - Create a new map
3. In the third exercise, you reproject layers to a shared SRS.
4. In the fourth (and optional) exercise, you create a map of Denmark using online data.

#### Required R libraries

We will use the sf, raster, and tmap packages.
Additionally, we will use the spData and spDataLarge packages that provide new datasets. To install the spDataLarge please load it directly from the author's github repository at https://github.com/Nowosad/spDataLarge. If the line in the chunk below does not work, check the repository for the other options.

```{r}
library(sf)
library(raster)
library(tmap)
library(spData)

# To install the spDataLarge uncomment the next line
install.packages("spDataLarge", repos = "https://geocompr.r-universe.dev")
library(spDataLarge)
```


#  Exercise 1 - Make a map

In this exercise you will learn to make a map with tmaps library, by adding spatial data layers and modifying the arguments that specify their rendering

#### Data sets  

We will use two data sets: `nz_elev` and `nz`. They are contained by the libraries
The first one is an elevation raster object for the New Zealand area, and the second one is an sf object with polygons representing the 16 regions of New Zealand.

#### Existing code

Here is the code to create a new map of New Zealand (following Robin Lovelace's example).Your role is to improve this map based on the suggestions below.

```{r initial-map}
tm_shape(nz_elev)  +
  tm_raster(title = "elev", 
            style = "cont",
            palette = "BuGn") +
  tm_shape(nz) +
  tm_borders(col = "red", 
             lwd = 3) +
  tm_scale_bar(breaks = c(0, 100, 200),
               text.size = 1) +
  tm_compass(position = c("LEFT", "center"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "A. Sobotkova, 2022") +
  tm_layout(main.title = "My map",
            bg.color = "orange",
            inner.margins = c(0, 0, 0, 0))
```


## Tasks 

1. Change the map title from "My map" to "New Zealand".
2. Update the map credits with your own name and today's date.
3. Change the color palette to "-RdYlGn". 
  (You can also try other palettes from http://colorbrewer2.org/)
4. Put the north arrow in the top right corner of the map.
5. Improve the legend title by adding the used units (m asl).
6. Increase the number of breaks in the scale bar.
7. Change the borders' color of the New Zealand's regions to black. 
  Decrease the line width.
8. Change the background color to any color of your choice.


## Your solution

```{r improved-map}
tm_shape(nz_elev)  +
  tm_raster(title = "elev in m asl", 
            style = "cont",
            palette = "-RdYlGn") +
  tm_shape(nz) +
  tm_borders(col = "black", 
             lwd = 1.5) +
  tm_scale_bar(breaks = c(0, 50, 100, 150, 200),
               text.size = 1) +
  tm_compass(position = c("RIGHT", "top"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "N. Nygaard, 09/02/2022") +
  tm_layout(main.title = "New Zealand",
            bg.color = "blue",
            inner.margins = c(0, 0, 0, 0))
```


# Exercise 2 - Inspect and display spatial data

We will use two data sets: `srtm` and `zion` , explore them and make a map.
The first one is an elevation raster object for the Zion National Park area in Israel, and the second one is an sf object with polygons representing borders of the Zion National Park.

Read two new datasets, `srtm` and `zion`, using the code below and then follow the instructions and answer the questions related.

####  Preparation: Load data from provided package
You need to have installed the spDataLarge package for the following code to work. 

```{r load-data}
install.packages("rgdal")
srtm  <-  raster(system.file("raster/srtm.tif", package = "spDataLarge"))
zion  <-  read_sf(system.file("vector/zion.gpkg", package = "spDataLarge"))
```

## Instructions and questions
Type answers to the questions as code comments next to or under the code used

* Display the `zion` object and view its structure.

  - What can you say about the content of this file?
  - What type of data does it store? 
  - What is the coordinate system used?
  - How many attributes does it contain?
  - What is its geometry?

* Display the `srtm` object and view its structure.

  - What can you say about the content of this file? 
  - What type of data does it store?
  - What is the coordinate system used? 
  - How many attributes does it contain?
  - How many dimensions does it have? 
  - What is the data resolution?
 
## Your solution and answers

```{r inspect-data}



#using the head function to inspect the zion file:
head(zion)

#This tells me that the file contains 1 feature and 11 fields, and furthermore that it is a polygon with registered max and min x and y-axis values. This indicates that it is a surface object, eg. a building, island or something similar. It further tells me that it uses a projected coordinate system and that we are in the northern hemisphere. It also seems that it it measured in meters rather than degrees. 

#Now that I know the file is a vector, I can use the st_crs function to find more information of the coordinate system of the object. 

st_crs(zion)

#This indicates that the coordinate system is using GRS80 (NAD83), which means it is an ellipsoid. It further affirms that it is measured in metres and not longitude and latitude, which also means it is a projected CRS, projected onto a 2D surface with northing and easting. It furhter has the ESPN ID 8611. 



# In order to inspect the srtm file, I will be using functions from the raster library, as it is a raster object.



# first using the extent function to get the extent of the srtm file:
extent(srtm)
#The file is a raster object containing coordinates ranging from -113.2396 to -112.8521 on the x-axis and from 37.13208 to 37.51292 on the y-axis, indicating that we are in the north-western hemisphere, more specifically in the Apple valley in USA and the area surrounding this place, ranging from the mentioned coordinates that are measured in degrees.

#Now using the crs function to find out what coordinate system is used:
crs(srtm)
#This indicates that it is a longitude/latitude datum (WGS84) coordinate system. This means that it is an unprojected coordinate system, and thereby a geographic. It has the EPSG ID 9122

#using the ncell function to find out how many grid cells the srtm file contains
# MANGLER: How many attributes does it contain?
#- How many dimensions does it have? 
#- What is the data resolution?
ncell(srtm)


```


# Exercise 3 - Reconciling Coordinate systems

Now that you know that coordinate systems differ, make them compatible!


## Instructions

* Reproject the `srtm` dataset into the coordinate reference system used in the `zion` object. 

  - Create a new object `srtm2`
  - Visualize the results using the `plot()` function.

* Reproject the `zion` dataset into the coordinate reference system used in the `srtm` object.
  - Create a new object `zion2`
  - Visualize the results (`zion2` together with `srtm2`) using the `plot()` function.

## Your solution

```{r}

library(raster)

#In order to get the CRS from the srtm object, I use the `crs()` function with argument `asText = TRUE` to get the CRS as a string and save this as `the_crs`.

the_crs <- crs(srtm, asText = TRUE)

#I then use the `st_transform()` to transform the zion object to the CRS in `the_crs`, and save it in a new object "zion2"

zion2 <- st_transform(zion, crs = the_crs)

#I then use the head function, to check if the CRS's match 

head(zion2)


#Now doing the exact same thing - converting the CRS for the srtm to the same as the one from zion, and saving it in a new object "srtm2"

the_other_crs <- crs(zion, asText = TRUE)

srtm2 <- projectRaster(srtm, crs = the_other_crs, method = "ngb")


#now I visualize the two maps together with the raster in the background and the polygon on top, using the plot-function:


plot(srtm2)
plot(zion, add=TRUE,
     col="transparent")

```



